---
title: "Parsing XML using chunked_invoke_rows"
author: "Curtis Alexander"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Parsing XML using chunked_invoke_rows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction
A common use of the `chunked_invoke_rows` function is to iteratively parse XML that is stored in a column within a database table.  This pattern can be applied to any large query result set that requires a custom function to be applied to each row.

## Parsing XML
The example below details how one would use the `xml_extract` function along with `chunked_invoke_rows` to iteratively parse XML.  Obviously the result set below will fit entirely into memory and doesn't need to be chunked.  Again, this is to demonstrate a pattern than can be used with much larger result sets.

The example below requires the `DBI`, `dplyr`, and `RSQLite`, packages.  The function `xml_extract` requires `purrr` and `xml2`.

### Prepare Database
The code below prepares an `RSQLite` in-memory database based on the `mtcars` dataset.  The code also will create a new column, `xml_col`, that will be used to demonstrate XML parsing.

```{r, eval = TRUE, echo = TRUE}
library("CRAmisc")
suppressMessages(library("dplyr"))

# function to create an XML document
xml_create <- function(car, hp, cyl) {
  xml_doc <- paste0("<root><car>",
                    car,
                    "</car><stats hp=\"",
                    hp,
                    "\" cyl=\"",
                    cyl,
                    "\"></stats></root>")
  gsub("[\r\n]", "", xml_doc)
}

mtcars_with_xml <- mtcars %>%
  mutate(car = attr(., "row.names"),
         xml_col = xml_create(car = car,
                              hp = hp,
                              cyl = cyl)) %>%
  select(car, mpg, hp, cyl, xml_col)  # only keep 5 columns

# in memory SQLite database
dbcon <- DBI::dbConnect(RSQLite::SQLite(), dbname = ":memory:")
invisible(DBI::dbWriteTable(dbcon, "mtcars_with_xml", mtcars_with_xml))
```

### Create Query and Send to Database
Next, create a query and send to the database.  The function `dbSendQuery` will submit a synchronous query to be run on the database and will block until it has completed.  But the result set is left on the database and is not brought back until `DBI::dbFetch` is called.

```{r, eval = TRUE, echo = TRUE}
sql_query <- "SELECT * FROM mtcars_with_xml where hp > 100"
res <- DBI::dbSendQuery(dbcon, sql_query)
```

### Write Custom Parsing Function
The function that is passed to `chunked_invoke_rows` should take in a dataframe and return a transformed dataframe.  The function below makes use of the `xml_extract` function which is just a convenience wrapper around the parsing functions found in the `xml2` package.  In addition, `xml_extract` makes use of `purrr::safely` to silence parsing errors.  This is useful when working with large result sets.

```{r, eval = TRUE, echo = TRUE}
# f should accept a dataframe and return a dataframe
f <- function(df) {
  df %>%
    dplyr::rowwise() %>%
    dplyr::mutate(
      xml_car = xml_extract(
        x = xml_col,
        xpath = "//root/car",
        extract_type = "text"
      ),
      xml_hp = xml_extract(
        x = xml_col,
        xpath = "//root/stats",
        extract_type = "attr",
        extract_value = "hp"
      ),
      xml_cyl = xml_extract(
        x = xml_col,
        xpath = "//root/stats",
        extract_type = "attr",
        extract_value = "cyl"
      )
    )
}
```

### Iterate Over Chunks
Invoke `chunked_invoke_rows` to iteratively parse the XML document within each row of the dataframe.

```{r, eval = TRUE , echo = TRUE}
# process 5 rows at a time
mtcars_with_xml_parsed <- chunked_invoke_rows(res, f, 5) %>%
  dplyr::select(-hp, -mpg, -cyl)
knitr::kable(mtcars_with_xml_parsed, format = "html", padding = 2)
```

### Cleanup
Close the database connection.  The result set does not need to be cleared (`DBI::dbClearResult(res)`) as that occurs within `chunked_invoke_rows`.

```{r, eval = TRUE, echo = TRUE}
DBI::dbDisconnect(dbcon)
```
